---
title: "Practical Machine Learning Project"
author: "Jiang Jin"
date: "October 24, 2015"
output: html_document
---

# Introduction    
Modern portable devices like Jawbone Up, Nike FuelBand and Fitbit help people collecting a large amount of data about activity. A group of enthusiasts who take measurements about themselves regularly to improve their health, many of them trends to focus on how much of a particular activity they do, instead of how well they do these activities.         

Data here collected accelerometers on the belt, forarm, arm and dumbell when the testers were asked to perform barbell lifts correctly and incorrectly in 5 different ways.    
The project here is to use data perform prediction of how they did the exercise.

Data preprocessing:     
1. Load Data
```{r}
setwd ("~/Desktop/WorkDirectory/Practical-Machine-Learning-Project")
library(caret)
library(rpart)
library(randomForest)
# load data
training <- read.csv("./data/pml-training.csv", row.names=1)
dim (training)

testing <- read.csv("./data/pml-testing.csv", row.names=1)
dim (testing)

```
Training and testing data set had 159 variables, training data had 19622 observations while testing data had 20 observations.

2. Data cleaning
Here we remove the missing values and null values.
```{r}
sum(complete.cases(training))
training1 <- training[,colSums (is.na(training))==0]
testing1 <- testing [,colSums (is.na(testing))==0]
dim (training1); dim (testing1)

# remove zero value
nsv <- nearZeroVar(training1, saveMetrics = TRUE)
train2 <- training1[,!nsv$nzv]
dim(train2)

nsv1 <- nearZeroVar (testing1, saveMetrics = TRUE)
test2 <- testing1[, !nsv1$nzv]
dim(test2)

```
After data cleaning, training set contains 19622 observations and 58 variables, while testing set contains 20 observations and 58 variables.


3. Data Splitting
Split training set into training and testing, 60% for training, 40% for cross validation.

```{r}

inTrain <- createDataPartition(y=train2$classe, p=0.6, list=FALSE)
train <- train2 [inTrain,]
crossValidation <- train2[-inTrain,]

dim(train)
dim(crossValidation)

```
4. Modeling
Use training part do modeling. 
"classe" is the factor to define activity type.

```{r}
library(caret)
modFit <- train(classe ~., method="rf", data=train, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE )
print (modFit$finalModel)

# library(rattle)
# fancyRpartPlot (modFit$finalModel)

```

5. Accuracy
Compute accuracy in training set and cross validation set.

```{r}
# train set
trainPre <- predict(modFit, train)
Matrix1 <- confusionMatrix (trainPre, train$classe)
Matrix1
# cross validation set
cvPre <- predict(modFit, crossValidation)
Matrix2 <- confusionMatrix (cvPre, crossValidation$classe)
Matrix2
```
From the tet we can see the accuracy in training set is 1, P-value <0.05. 
Accuracy in cross validation set is 0.997, P-value <0.05.

6. Test's prediction
Here we use modFit to predict testing data set.

```{r}
testPre <- predict (modFit, test2)
testPre
summary(testPre)

```
According to result, the 20 cases, 7 are class A, 8 are class B, 1 is class C, 1 is class D, 3 are class E.







